Computational hysteresis:\\ inter-task effects in human computation

### 2014-10-22 docs/publications/2014.09_SCIENCE/man.tex
work on paper.  Introduce new data analysis and plots

\begin{sciabstract}
Microtask platforms offer a fast, flexble, and inexpensive source of labour for
clerical tasks \cite{Finnerty2013, chandler2013breaking, Berinsky2012351} and 
recently have been adopted by researchers to augment datasets and solicit 
participants \cite{paolacci2010running, Berinsky2012351, chandler2013breaking}.
Some computer scientists describe such platforms as a new computing 
architecture, and seek to characterize “human co-processing units” (HPUs) in 
analogy to CPUs and GPUs \cite{5543192}. However, it is known that people are 
susceptible to 
priming effects\cite{No2007,Swaab200299,Gopher2000308,sohn2001task,Ghuman17062008,BJOP1796,BJOP1826,Gass1999549}, which stands to complicate the HPU 
perspective.  Here, we introduce a framework to recast this psychological 
phenomenon as a kind of computational hysteresis. 
We apply this framework to study \textit{inter-task effects} that arise when 
microtask 
workers perform a sequence of tasks. Surprisingly, earlier tasks can have a 
very strong influence on workers during later tasks---stronger than overtly
framing the tasks as serving different research goals.   
This result has major implications for the design of crowdsourcing initiatives.
The framework we introduce provides a general measure of priming severity, 
and, in combination with machine learning tools, enables measurement without
knowing ahead of time how priming will affect output.
\end{sciabstract}

@@

The task requester can interact with the platform like a compute server, 
seamlessly 
integrating human and machine computation.  Researchers coined the 
term HPU (Human co-Processing Unit), viewing the introduction 
of microtask platforms as a new computing architecture\cite{5543192}.  
Attempts have been made to formalize the HPU instruction set, and libraries 
that provide high-level APIs (Application Programming Interfaces) for writing 
algorithms involving human computation have been developed \cite{little2010turkit,minder2011crowdlang,minder2012crowdlang,kittur2011crowdforge}.

Here, we highlight an important way in which HPUs differ from CPUs, with 
serious implications for any crowdsourcing initiative.  Common sense would 
dictate that people's output during a task will be stochastic: when the same
task is given to different people, or given to the same person at different 
times,
the outputs may, in general, differ.  Stochasticity is not itself problematic,
since the analysis of probabilistic algorithms is fairly 
well-understood. But other non-random effects may also bear on HPU output. 
It is known that people are susceptible to priming effects 
\cite{BJOP1796,No2007,beller1971priming}, and, in particular, task-repetition 
effects \cite{Gass1999549,sohn2001task}.

Such repetition effects, or
\textit{inter-task effects}, are neither random, nor are they formally part 
of the instructions given to the HPU.  The ``leakage'' of information from 
recently performed tasks into HPU outputs is a potentially strong source of 
systematic bias, which complicates the the algorithmic view of HPUs. 
Inter-task effects would amount to a kind of \textit{hysteresis}, meaning 
that HPU output is not only a function of the current input, but also of the 
history of inputs.

Here we make two contributions towards understanding inter-task effects
and HPU hysteresis in general.  We present a framework for treating this 
source of bias in the context of HPU algorithms, and we measure the impact of
inter-task effects experimentally in a canonical HPU task: the labeling of 
images.

Using the  Amazon Mechanical Turk (Mturk) platform, we solicited workers to 
perform a series of image-labeling tasks.  
In these tasks, workers labeled images featuring iconic cultural scenes,
food, table-settings, and various utensils and cookware. We divided the set of 
images shown to a given worker into an \textit{initial} and \textit{test} set, 
but, crucially, there was no apparent distinction between these sets from the
perspective of the worker.  We varied the initial set of images, while 
keeping the test set the same, to analyze the effect that that the initial 
images had on the labels that workers attributed to the test set.

There has been considerable investigation into the factors that affect the 
quality and nature of microtask work.  These include such factors as the 
level of 
pay \cite{kazai2013analysis}, training \cite{le2010ensuring}, pre-screening of 
workers \cite{paolacci2010running}, and user-interface design 
\cite{Finnerty2013}.  Researchers have also investigated \textit{framing}, 
by testing the effects of describing the workflow context 
\cite{Kinnaird2012281}, the purpose of tasks 
\cite{chandler2013breaking}, or using an alternative problem discription
\cite{thibodeau2013natural}.  To our knowledge, no study has investigated 
inter-task effects on microtask platforms.

As a point of comparison, we included treatments in which we induce 
\textit{framing}.
Before the workers label images, we indicate that the work 
was funded by a research lab having a particular area of interest, to see the 
influence this has on the workers’ subsequent labels. 
Surprisingly, inter-task effects were stronger than framing, in some cases 
about twice as strong. 
Our results show that initial tasks can significantly alter the subject 
matter on which worker's labels focus. We also found that inter-task 
effects can induce workers to provide labels of greater or lesser specificity. 
This suggests that careful consideration should be given to the bundling of 
tasks when designing a study using a microtask platform.

To characterize inter-task effects, we present a novel, generalized definition 
of \textit{HPU priming}. Priming is traditionally used to elucidate the
internal psychological mechanisms of cognition.  But here we are concerned with
understanding the outward effects that priming phenomena have on HPU 
algorithms. Thus, before we discuss inter-task effects experimentally, we 
discuss formal definitions for priming as computational phenomenon.  Our
discussion will lead to a natural definition of priming that enables 
comparisons of the severity priming effects accross applications, and enables
the measurement of priming using machine learning tools.

\paragraph*{Priming as computational hysteresis.}
In psychology, priming effects are studied because they hint at the 
psychological mechanisms at play during certain cognitive tasks. But in the 
context of human computation, the focus should be on the outward consequences: 
the potential for algorithms to execute differently when ``run'' on differently
primed HPUs.

We will present three definitions of priming with this practical concern in 
mind, and then describe how to measure priming in practice. To proceed with 
the definitions, we will assume that two populations of HPU’s, $\mathcal{P}$ 
and $\mathcal{Q}$ were 
sampled randomly from the same initial population $\mathcal{R}$, and then were 
exposed to (possibly) different priming treatments. The HPUs in $\mathcal{P}$ 
and $\mathcal{Q}$ perform some 
task, leading to popualtions of work-products, $P$ and $Q$.

For the first defenition, let us address the concern that differently primed 
HPUs may produce ``biased'' work-products.

\begin{mydef}
	{\upshape Maximum Bias $\theta_\mathrm{bias}$:}
	Let $\mathcal{D}$ be a binary decision algorithm, which takes a 
	work-product
	$r \in P \cup Q$ as input, and outputs a bit $\mathcal{D}(r)=b$, where 
	$b \in \{0,1\}$. Denote the \emph{bias} of $\mathcal{D}$, due to the
	different priming of $\mathcal{P}$ and $\mathcal{Q}$ by
	$\theta^\mathcal{D}_\mathrm{bias}(P,Q)$ and define it to be the difference 
	in the expected output of $\mathcal{D}$ when acting on work-products from 
	$P$ vs $Q$:
	$$
	\theta^\mathcal{D}_\mathrm{bias}(P,Q) = 
		\left| 
			\mathrm{E}\left\{ \mathcal{D}(p) \right\} 
			- \mathrm{E}\left\{ \mathcal{D}(q) \right\} 
		\right|,
	$$
	where $p \in P$, $q \in Q$.  Then, define the difference in priming 
	between $P$ and $Q$ to be the worst-case bias, i.e. the bias of the most
	susceptible decision:
	$$
	\theta_\mathrm{bias} = 
		\sup_\mathcal{D} \theta^\mathcal{D}_\mathrm{bias}(P,Q)
	$$
\end{mydef}

Thus, $\theta_\mathrm{bias}$ has a straightforward interpretation as an upper 
bound on the amount of bias induced on any yes-or-no decision, due to the 
difference in priming of $\mathcal{P}$ and $\mathcal{Q}$.

For the next definition, we attempt to address directly the extent to which 
two populations of work-products are ``different'', by supposing that they are 
different to the extent that they can be distinguished algorithmically. To 
facilitate this definition, we make use of the notions of a 
\textit{distinguisher} and a \textit{validation} test.

\begin{mydef}
	{\upshape Algorithmic distinctness $\theta_\mathrm{dist}$:}
	Let $\mathcal{D}$ be a distinguisher algorithm, which takes work-product 
	$r \in P \cup Q$ as input, and outputs a \emph{guess} 
	$\mathcal{D}(r) \in \{0, 1\}$. Now define a validation test 
	$V(\mathcal{D}, R_0, R_1)$ as follows. A bit is drawn uniformly at random, 
	$b \in_R \{0, 1\}$, and then a work-product $r$ is chosen uniformly at 
	random from one of the populations, decided by the bit: $r \in_R R_b$. 
	Feed $r$ to $\mathcal{D}$, and if $\mathcal{D}(r) = b$, we say 
	$\mathcal{D}$ guessed correctly, and the value of $V$ is 1, 
	otherwise V = 0.

	A distinguisher that guesses randomly will still be correct half of the 
	time. Therefore, denote the performance of $\mathcal{D}$ as 
	$\eta_\mathcal{D}(P, Q)$, and define:
	$$ 
		\theta_\mathrm{dist} = \sup_\mathcal{D} \eta_\mathcal{D}(P,Q)
	$$

\end{mydef}
In the above definition, the performance simply rescales accuracy, so that when
accuracy is 1/2 (i.e. as good as chance) performance is 0, and when accuracy 
is perfect, performance is 1. This definition casts the measurement of priming 
effects in terms of a binary classification problem, which is well-studied in 
the field of machine learning.

The two definitions we have already mentioned are, in fact, equivalent, and 
correspond to a third, more fundamental formulation known as the total 
\textit{variational distance}, or $l_1$-distance, or simply the “statistical 
distance” ($D_{l_1}$):
$$
	D_{l_1}(P,Q) \equiv \frac{1}{2} \sum_{x \in \mathcal{X}} 
	\left| 
		f_P(x) - f_Q(x)
	\right|,
$$
where $f_P(x)$ denotes the probability of observing x when sampling from P. 
This leads us to our third definition:

\begin{mydef}
	{\upshape Priming as $l_1$-distance:}
	The difference in priming of $\mathcal{P}$ and $\mathcal{Q}$ is the 
	$l_1$-distance between the distributions of their work-products, 
	$D_{l_1}(P,Q)$.
\end{mydef}

We provide proofs of the equivalence between these definitions in the 
supplementary material. Despite their equivalence, we show the three 
definitions because they offer different theoretical, practical, and 
interpretive entry points. Definition 1 provides a practical motivation, and 
the straighforward interpretation of priming as bias. Definition 3 draws a 
clear link to existing statistical theory. And Definition 2 introduces machine 
learning tools, which allow us to gain traction when the purely statistical 
tools fail.

\paragraph{Measuring priming.}
Care must be taken in estimating $D_{l_1}$ from samples so as not to obtain
an inflated estimate. In the so-called naive approach, one uses the frequency 
with which a given work-product $x$ is obtained from HPUs in $\mathcal{P}$ to 
estimate the probability mass function at $x$; 
$\hat{f}_P(x) = (\sum_{X\in P} \mathbf{1}_{[X=x]})/|P|$.
One can then estimate 
$D_{l_1} = \sum_{x \in \mathcal{X}} \left|\hat{f}_P(x) - \hat{f}_Q(x) \right|/2$ using only the sample frequencies. But the number of samples, $|P|$ and 
$|Q|$, must be much larger than the number of possible work-outputs, 
$|\mathcal{X}|$, otherwise, this method produces an enormously inflated 
estimate.

Recently, attempts have been made to provide better estimators. Valiant, 
provides a canonical estimation approach based on first estimating the unseen 
region of the support. This algorithm appears to perform very well in practice.
Batu provided an alternative approach, which bypasses estimating the 
probability mass functions, and relies on the more fundamental concept of 
\textit{collisions}.

Both methods converge much more quickly than the naive approach. However, two 
limitations remain that prevent their application. First, these relatively 
recent approaches are not accompanied by any absolute guarantees of 
statistical confidence, so it is not yet clear how to build confidence 
intervals for a given estimate. Second, sophisticated as these methods are, 
they still cannot provide reliable estimates when none of the elements of the 
support recur in the samples (that is, when we never observe the same 
work-product twice in our sample sets).

It may seem unreasonable to hope to measure $D_{l_1}$ when our samples are so 
megre that no specific work-product recurrs, but this level of sparsity is 
routinely faced in classification tasks. As a concrete example, a facial 
recognition 
system (which classifies images as face or non- face) is expected to work on 
images that have never been seen before, and its design cannot be predicated on
covering a significant fraction of all possible images during training. In our 
experiment, we demonstrate that the quintessential microtask, image labeling, 
is one for which the measurement of priming effects can be approached using 
classifiers.

\paragraph{Experimental setup.}

image labeling is common\ldots
\cite{chandler2013breaking,Berinsky2012351,Finnerty2013,paolacci2010running}.

We solicited 2300 MTurk workers to perform an image-labeling task using images 
that depicted meals, place-settings, and related items such as dishes, and 
kitchenware, and other artifacts naturally arising in dining contexts. Workers 
were randomly assigned into one of four treatments, described below. In all 
treatments, workers were given brief instruc- tions before labeling any images.
In some treatments, workers were then also shown a \textit{framing slide}, 
indicating 
the tasks were funded by a group having a particular research focus. The tasks 
consisted of single images, and workers were asked to provide five descriptive 
labels for each. The last five images that workers labeled, the 
\textit{test set}, were the same for all treatments.

The first two treatments, IMG:OBJ and IMG:FOOD, tested inter-task effects. 
Workers in these treatments labelled ten images. In the IMG:OBJ treatment, the 
\textit{initial set} of five images labeled by workers contained objects such 
as utensils, kitchenware, and table-settings, but no food items. In the 
\textsc{img:food}, the initial set contained centered close-ups of meals, 
without
any surrounding objects (except sometimes the dish supporting the food). No 
interruption or distinction made between the initial and test sets from the 
view of the workers. To study the positional dependance of inter-task effects,
these treatments were subdivided, and workers were shown one of five 
permutations. There are of course 5! = 120 permutations of five images, but 
investigating all permutations was too expensive, and probably redundant, so 
we chose 5 permutations that ensure each image occupies each position within 
the test set. We discuss the benefits and drawbacks of this choice in the 
supplementary material.

Workers in the treatments \textsc{frm:food} and \textsc{frm:obj} were shown a 
framing slide, which bore a message reading: ``Funded by the laboratory for 
the visual perception of \{Food and Ingredients $\vert$ Objects and Tools\}''. 
After the framing slide, workers in these treatments proceeded directly to 
the test set (and so labeled only 5 images).

\paragraph{The strength of inter-task effects}
We tested for inter-task effects by training a Naive Bayes classifier to 
classify work-products among IMG:OBJ and IMG:FOOD, using only the labels 
attributed to the test set of images. Based on Definition 2, no classifier 
can perform better than the actual difference in priming, so the performance 
of Naive Bayes classifier provides a lower bound on priming difference. We 
used leave-one-out validation to estimate the classifier per- formance. We 
tested for framing effects in a similar way, by training a Naive Bayes 
classifier to classify work-products among \textsc{frm:food} and 
\textsc{frm:obj}.

Overall, the first image in the test set showed the strongest signal for 
priming effects. Fig. 3 shows the empirical lower bound for difference in 
priming due to inter-task effects and framing based on the classifier 
performance when using labels from the first test image. Statistically 
significant levels of priming were detected for both modalities, but the 
inter-task effects were stronger (hypothesis test). This suggests that strong 
bias can result solely from the sequence of tasks performed by a worker, even 
when those tasks are of the same type.

In contrast, the classifier never performed significantly better than chance 
when distinguish- ing treatments that differed only in framing. Thus, framing 
effects were rather weak. In fact, framing effects were generally weak in our 
results, so we omit treatments involving framing from the remainder of our 
presentation to save space. We encourage the interested reader to look at the 
results for these treatments in the supplementary material.

\paragraph{Inter-task effects and the orientation of focus.}
Since the initial images were chosen to emphasize either food (ingredients 
set) or culture (cultural set), we looked for effects on the number of 
culture- and food-oriented labels that workers attributed to the test image 
set.

To this end, we constructed an ontology from the labels attributed to the test 
images. In the ontology, edges point from more general labels to more 
specific ones. For example, the ontology contains the path \texttt{food} $\to$ 
\texttt{ingredients} $\to$ \texttt{vegetables} $\to$ \texttt{tomato}.

Since food is a central feature of culture, our ontology contains many labels 
that have both food and culture in their ancestry. Nevertheless, there were 
many food-oriented labels, such as bread, which lacked specific cultural 
connections, as well as non-food, culture- orientedlabels,suchasrussian dolls.

When we tallied labels attributed to the first image of the test set, we found 
that workers from CULTimg produced significantly more culture-oriented labels 
and less food-oriented ones than those from AMBG (see Fig. 2A). The inter-task 
effect was so strong that the proportion of food- and culture-oriented labels 
in CULTimg was essentially the reverse of that in AMBG. However, the 
composition of labels attributed by INGRimg was not significantly different 
from those attributed by AMBG. This shows that inter-task effects can 
profoundly alter workers' focus. Again, we will argue later that the ambiguous 
and ingredients image sets are, in fact, quite similar.

>>>>

\begin{sciabstract}

People still outperform sophisticated artificial 
intelligince systems in many tasks requiring general knowledge, 
qualitative judgment, and visuo-spatial reasoning. Recently, microtask 
platforms enable the programmer to import human judgment directly into 
algorithms: programs can post tasks, to be completed by users of such 
platforms in near real-time.
Some computer scientists consider microtask platforms to be a fundamentally
new computing architecture, but psychological effects can introduce systematic
bias. 
Here we investigate the effects that earlier tasks have on the results of 
subsequent ones, which we call \textit{inter-task effects}.
In a canonical image labeling task, we found that inter-task effects are
very strong---stronger than overtly framing tasks in terms of 
particular research goals.  However, in some cases, inter-task effects can
be leveraged to elicit more nuanced responses.
To study inter-task effects we introduce a framework that characterizes HPU 
bias in an application-independant way, based on its potential to alter an 
algorithm's path of execution.  Our findings suggest that careful consideration
should be given to inter-task effects when designing crowdsourcing studies.

%microtask worker performs a series of tasks.  We find that early tasks can
%have a strong effect on later tasks, and that these effects can be stronger
%than those brought about by framing.
%
%course of human computation algorithms.
%application-independant quantity that measures the
%severity of priming.  We apply this framework to measure 
%
%in many
%tasks, especially ones requiring general knowledge about the world, qualitative
%judgment, and visuo-spatial reasoning.  Recently, microtask web platforms have 
%emerged. One can upload a batch of tasks to be completed by users who are paid 
%for their efforts.
%Though simple on the surface, these platforms dramatically enlarge the
%programmers toolkit: in effect they allow the programmer to import human 
%judgment into algorithms.  Some computer scientists consider
%such platforms to be a new computing architecture.  But people are very 
%different from electronic processors.  For example, context affects how a 
%person will respond to a prompt, and this complicates the formal treatment of 
%human computation.  To approach this issue, we introduce a framework 
%that quantifies the severity of priming in an application-independant way.
%We apply this technique to reveal an unforseen source of bias: the bias 
%induced by one task on another.  Surprisingly we find that 
%\textit{inter-task effects} introduce more bias than \textit{framing}, which
%has received more attention.
%
%more well-known 
%\textit{framing}, a more well-studied form of priming.
%
%Microtask platforms offer a fast, flexble, and inexpensive source of labour for
%clerical tasks \cite{Finnerty2013, chandler2013breaking, Berinsky2012351} and 
%recently have been adopted by researchers to augment datasets and solicit 
%participants \cite{paolacci2010running, Berinsky2012351, chandler2013breaking}.
%Computer scientists have described such platforms as a new computing 
%architecture, and seek to characterize “human co-processing units” (HPUs) in 
%analogy to CPUs and GPUs \cite{5543192}. However, it is known that people are 
%susceptible to 
%priming effects\cite{No2007,Swaab200299,Gopher2000308,sohn2001task,Ghuman17062008,BJOP1796,BJOP1826,Gass1999549}, which stands to complicate the HPU 
%perspective.  Here, we introduce a framework to recast this psychological 
%phenomenon as a kind of computational hysteresis. 
%We apply this framework to study \textit{inter-task effects} that arise when 
%microtask 
%workers perform a sequence of tasks. Surprisingly, earlier tasks can have a 
%very strong influence on workers during later tasks---stronger than overtly
%framing the tasks as serving different research goals.   
%This result has major implications for the design of crowdsourcing initiatives.
%The framework we introduce provides a general measure of priming severity, 
%and, in combination with machine learning tools, enables measurement without
%knowing ahead of time how priming will affect output.
\end{sciabstract}

@@

These platforms make it possible to seemlessly integrate human and 
machine computation.  Researchers coined the term HPU 
(Human co-Processing Unit) in analogy to CPUs and GPUs \cite{5543192}.  
Attempts have been made to establish an HPU instruction set.  For example,
the X library provides function calls such as 
\texttt{X.vote()}, and \texttt{X.map()}
 \cite{little2010turkit,minder2011crowdlang,minder2012crowdlang,kittur2011crowdforge}.

Of course, there are obvious differences between HPUs and their silicone
counterparts.  When facing the same task, people do not generally
respond in the same way.  Common sense would dictate that a certain amount
of variability in responses made by people is inescapable.  But an important 
part of the variation is bound to be systematic, due to psychological or 
demographic factors.

Here, we highlight an important source of bias with
serious implications for any crowdsourcing initiative. 
It is known that people are susceptible to priming effects 
\cite{BJOP1796,No2007,beller1971priming}, and, in particular, task-repetition 
effects \cite{Gass1999549,sohn2001task}.  Thus, a worker's response during
one task may depend in part on previous ones.  Such \textit{inter-task} effects
would amount to a kind of \textit{hysteresis}, meaning that HPU outputs are not
only a function of the current input, but also of the history of inputs.

There has been considerable investigation into the factors that affect the 
output from microtask work.  These include such factors as the 
level of 
pay \cite{kazai2013analysis}, training \cite{le2010ensuring}, pre-screening of 
workers \cite{paolacci2010running}, and user-interface design 
\cite{Finnerty2013}.  Researchers have also investigated \textit{framing}, 
by testing the effects of describing the workflow context 
\cite{Kinnaird2012281}, the purpose of tasks 
\cite{chandler2013breaking}, or using an alternative problem discription
\cite{thibodeau2013natural}.  To our knowledge, no study has investigated 
inter-task effects on microtask platforms.

We make two contributions towards understanding inter-task effects,
and systematic bias in HPUs in general.  First, we present a framework for 
quantifying bias in HPU outputs in terms of their potential impact 
algorithms.  This yields an approach for measuring the severity of bias 
that removes the need to understand the underlying psychological or 
demographic factors at play.  Second, we apply our framework to measure the
severity of inter-task effects on a canonical microtask: image labeling.

Using the  Amazon Mechanical Turk (Mturk) platform, we solicited workers to 
perform a series of image-labeling tasks.  We chose image labeling because it 
is one of the most common kinds of microtasks
\cite{chandler2013breaking,Berinsky2012351,Finnerty2013,paolacci2010running}, 
and it seems likely that this will remain the case in the long term because
of applications in computer vision research \cite{5543192}.
%In these tasks, workers labeled images featuring iconic cultural scenes,
%food, table-settings, and various utensils and cookware. 
We divided the tasks into an \textit{initial} and \textit{test} set, 
but, crucially, there was no distinction between these sets from the
perspective of the worker.  We varied the initial tasks, while 
keeping the test set the same, to analyze the effect of initial tasks on 
the labels provided during the test tasks.

As a point of comparison, we included treatments in which we induce 
\textit{framing}.
Before the workers labeled images, we indicated that the work 
was funded by a research lab with a particular area of interest, to see the 
influence this has on the workers’ subsequent labels. 
Surprisingly, inter-task effects were stronger than framing, in some cases 
about twice as strong. 
Our results demonstrate that initial tasks can change workers' focus in later 
tasks.  We also found that inter-task effects altered the level of specificity 
of worker's labels.  Our findings suggest that careful consideration should be 
given to the bundling of tasks when designing a study using a microtask 
platform.

\paragraph*{Computational hysteresis.}
Before describing our experiments, we dedicate some space to discussing
how we characterize and measure priming in HPU outputs.  Since HPU outputs are
stochastic, determining the effect of priming on an HPU population
amounts to determining the resulting change to the HPU population's 
distribution of outputs.

To refine this, we distinguish between methods that determine \textit{whether}
two random sources have the same distribution 
(e.g. Pearson's chi squared test), from those that determine
\textit{the extent} of that difference (e.g. one of the many measures of 
statistical divergence).  We are interested in measuring the extent of
difference because our practical concern is the potential for such
differences to alter the course of an algorithm.

To measure the effect of some priming treatment, we must of course draw a 
comparison to some standard population.  In many cases, there may be an 
obvious comparison between a primed and unprimed population.  However, we 
refrain from proposing a canonical neutrally-primed HPU population; what is 
considered ``unprimed'' in one scenario may be ``primed'' in another. 
Instead, we speak of two HPU populations being \textit{differently primed}.

We present three alternative, equivalent definitions for the 
\textit{difference in priming} of two HPU populations, $\mathcal{P}$, and 
$\mathcal{Q}$, which we denote $\theta(P,Q)$.  To set up the definitions,
we will assume that $\mathcal{P}$ and $\mathcal{Q}$ 
were sampled randomly from the same initial population $\mathcal{R}$, and 
then were exposed to (possibly) different priming treatments. Then, the HPUs 
in $\mathcal{P}$ and $\mathcal{Q}$ perform some task, leading to popualtions 
of work-products, $P$ and $Q$.

The first definition provides the motivation for our choice of divergence
measure, because it directly measures the proportion of the time that a
binary decision will be changed due to priming effects:

\begin{mydef}
	\label{def:bias}
	{\upshape Maximum Bias:}
	Let $\mathcal{D}$ be a binary decision function, which takes a 
	work-product
	$r \in P \cup Q$ as input, and outputs a bit $\mathcal{D}(r)=b$, where 
	$b \in \{0,1\}$. Denote the \emph{bias} of $\mathcal{D}$, due to the
	different priming of $\mathcal{P}$ and $\mathcal{Q}$ by
	$\theta_\mathcal{D}(P,Q)$ and define it to be the difference 
	in the expected output of $\mathcal{D}$ when acting on work-products from 
	$P$ and $Q$:
	$$
	\theta_\mathcal{D}(P,Q) = 
		\left| 
			\mathrm{E}\left\{ \mathcal{D}(p) \right\} 
			- \mathrm{E}\left\{ \mathcal{D}(q) \right\} 
		\right|,
	$$
	where $p \in P$, $q \in Q$.  Then, define the difference in priming 
	between $\mathcal{P}$ and $\mathcal{Q}$ to be the worst-case bias, i.e. 
	the bias of the most susceptible decision:
	$$
	\theta(P,Q) = 
		\sup_\mathcal{D} \theta_\mathcal{D}(P,Q)
	$$
\end{mydef}
If $\mathcal{D}$ is the condition of an \texttt{if}-statement, then the 
definition above is the change in the fraction of the time that a given 
path of execution is followed due to HPU priming.

While this is precisely what we want to measure, it is not at all clear how
to go about doing it.  The second formulation provides a practical connection
by rephrasing priming in terms of its distinguishability:

\begin{mydef}
	{\upshape Distinguishability:}
	Let $\mathcal{D}$ be a distinguisher algorithm, which takes work-product 
	$r \in P \cup Q$ as input, and outputs a \emph{guess} 
	$\mathcal{D}(r) \in \{0, 1\}$. Now define a validation test 
	$V(\mathcal{D}, R_0, R_1)$ as follows. A bit is drawn uniformly at random, 
	$b \in_R \{0, 1\}$, and then a work-product $r$ is chosen uniformly at 
	random from one of the populations, decided by the bit: $r \in_R R_b$. 
	Input $r$ into $\mathcal{D}$, and if $\mathcal{D}(r) = b$, we say 
	$\mathcal{D}$ \emph{guessed correctly}, and the value of $V$ is 1, 
	otherwise $V = 0$.

	A distinguisher that guesses randomly will still be correct half of the 
	time. Therefore, denote the performance of $\mathcal{D}$ by 
	$\eta_\mathcal{D}(P, Q)$, and let
	$$
		\eta_\mathcal{D}(P,Q) = 2\cdot \mathrm{E}\{V(\mathcal{D},P,Q)\} - 1.
	$$
	Then define the difference in priming between $\mathcal{P}$ and 
	$\mathcal{Q}$ to be the performance of the best possible distinguisher:
	$$ 
		\theta(P,Q) = \sup_\mathcal{D} \eta_\mathcal{D}(P,Q)
	$$
	\label{def:dist}
\end{mydef}

\textbf{Definition 2} relates the difference in priming to the 
\textit{optimal Bayes classifier} for determining whether a given HPU output
came from $P$ or $Q$.  This means that we can obtain a lower bound on
$\theta(P,Q)$ based on the performance of an actual classifier, which provides
recourse when purely statistical methods fail.

The statistical divergence measure underlying the previous definitions is the
divergence measure known as the ``total variational distance'' or 
$l_1$-distance:
\begin{mydef}
	\label{def:l1}
	{\upshape Total variational distance $\theta_{l_1}$:}
	Define the difference in priming between $\mathcal{P}$ and $\mathcal{Q}$ 
	to be equal to the total variational distance between the work-products 
	$P$ and $Q$:
	$$
	\theta_{l_1}
	= \sup_{A \subset \mathcal{X}} \left| \sum_{x \in A} f_P(x) - f_Q(x)\right|
	= \frac{1}{2} \sum_{x \in \mathcal{X}} \left| f_P(x) - f_Q(x) \right|,
	$$
	where $f_P(x)$ denotes the probability of observing x when sampling from P.
\end{mydef}
The equivalent formulation for $\theta(P,Q)$ in 
\textbf{definition \ref{def:l1}} provides a grounding in a fundamental
measure of statistical divergence, and a connection to information theory.  
The equivalence of the three definitions
follows from the known fact that no algorithm can have output that differs
on inputs $P$ and $Q$ more than $l_1 \times 100\%$ of the time.  We provide
proofs in the supplementary material.

The difficulty of measuring $\theta(P,Q)$ depends on what the universe of
possible outputs are, as well as how $P$ and $Q$ are actually distributed,
both of which are mainly determined by the task that the HPUs are asked to 
perform.  In the case where workers label an image with a free-text input, 
the number of different possible labels can be in the thousands or more.  If
the support, $\mathcal{X} = P \cup Q$, is large compared to the number of 
HPU outputs sampled, so that many labels appear only once (which in any case
is expected from Zipv's law), then estimating $\theta$ directly from the
observed frequencies of labels is difficult.

One approach to estimating the $l_1$-distance from samples, which appears
reasonable on the surface, is to use the observed frequencies of each output
$x$ from $P$ and $Q$ to estimate $f_P(x)$ and $f_Q(x)$, and calculate 
$\theta(P,Q)$ directly from \textbf{definition 3}.  This is the so-called naive
approach to estimating $l_1$-distances, and is aptly named because it tends
to produce very inflated values.  More recent techniques have for estimating
$l_1$ have been suggested by Valiant \cite{val-thesis}, and Batu 
\cite{batu2000testing,batu2013testing}.  We reserve the discission of these
techniques to the supplementary material; they do not turn out to be useful
because they do not provide confidence intervals.  Thus our only recourse is
to test the performance of classifiers at distinquishing $p \in P$ from 
$q \in Q$, which allows us to estimate a lower bound on $\theta(P,Q)$.  This 
turns out to be sufficient, and has the advantage that, so long as they are
learnable, it is not necessary to understand exactly how $P$ and $Q$ differ.

\paragraph{Experimental setup.}
We performed two experiments, soliciting respectively 900 and 2300 MTurk 
workers to label images depicting iconic cultural scenes, meals, ingredients, 
table-settings, and various utensils and cookware. In these experiments,
workers were randomly assigned to treatments, and depending on this 
assignment, were exposed to different framing primes, given a different 
set of initial images, and shown a different permutation of the test images. 
For both experiments, workers were first shown brief instructions.  
The framing prime was then shown, unless the given 
treatment involved no framing.  Workers then performed a series of 
image-labeling tasks and had to provide five labels per image. 
For the purpose of analysis, we denote the first five images as the 
\textit{initial set}, and the last five as the \textit{test set}, but no
distingction was made from
the perspective of the worker.
Table 1 sumarizes the various experimental treatments for both 
experiments.

\begin{table}[t]
\centering
	\begin{tabular}{ l  l  l }
		\hline                       
		Treatment & Framing prime & Initial image set	\\ 
		\hline                       
		\textit{Experiment 1}\\
		$\textsc{ambg}$ & None & Ambiguous\\
		$\textsc{img:cult}$ & None & Cultural\\
		$\textsc{img:ingr}$ & None & Ingredients\\
		$\textsc{frm:cult}$ & Cultural & Ambiguous\\
		$\textsc{frm:ingr}$ & Ingredients & Ambiguous\\
		\hline  
		\textit{Experiment 2}\\
		$\textsc{img:food}$ & None & Food\\
		$\textsc{img:obj}$ & None & Objects\\
		$\textsc{frm:food}$ & Food & None\\
		$\textsc{frm:obj}$ & Objects & None\\
		\hline  
	\end{tabular}

	\caption{ \footnotesize{ 
		Workers were uniformly randomly assigned to one of the 
		treatments listed above. 
		The full funder names used were 
		Cultural: ``The Global Foundation for Cultural Recognition'' and 
		Ingredients: ``The National Foundation for Nutritional Awareness''.  
		The ambiguous, cultural, and ingredients initial image sets are shown 
		in \textbf{Figs. 2}, \textbf{3}, and \textbf{4}.
	}}
	\label{table:1}
\end{table}

\begin{figure}
	\centering
	\includegraphics[scale=0.7]{figs/tasks.pdf}
	\caption{Examples of slides used for the image labeling tasks. A) instructions shown to all workers; B) a framing slide; C) a slide from the initial set for treatment IMG:FOOD; D) a slide from the initial set for treatment IMG:OBJ; E) a slide from the test set shown in all treatments. The full set of slides used for all treatments is presented in the supplementary material.}
	\label{fig:task}
\end{figure}

\paragraph{The strength of inter-task effects.}
\begin{figure}
	\centering
	\includegraphics[scale=1]{figs/theta.pdf}
	\caption{
		Difference in priming observed for various pairs of worker populations.
		The pairs of populations differed in terms of the initial tasks they
		performed (bars labeled by \textit{inter-t.}) or in terms of the framing
		to which they were exposed (bars labeled by \textit{frame}; see table 1 for
		details about the treatments).  The difference in priming was 
		measured based on the accuracy of a naive Bayes classifier in 
		resolving which treatment workers belonged to, as described in the 
		text.  
	}
	\label{fig:theta}
\end{figure}

%\paragraph{Dynamics of inter-task effects}
%\begin{figure}
%	\centering
%	\includegraphics[scale=0.7]{figs/theta_by_pos_img.pdf}
%	\caption{
%		Difference in priming observed for various pairs of worker populations.
%		The pairs of populations differed in terms of the initial tasks they
%		performed (bars labeled by ``inter-t.'') or in terms of the framing
%		to which they were exposed (bars labeled by ``frame''; see table 1 for
%		details about the treatments).  The difference in priming was 
%		measured based on the accuracy of a naive Bayes classifier in 
%		resolving which treatment workers belonged to, as described in the 
%		text.  
%	}
%	\label{fig:theta_pos_img}
%\end{figure}

We analyzed the overall strength of inter-task effects and framing, by looking
at pairs of treatments that performed the same test tasks, but had done
different initial tasks or had been exposed to different framing.
For each such pair, we assessed the performance of classifier
in predicting which treatment workers belonged to, based on the labels they
provided during the test tasks (see fig 2).  
Based on \textbf{Definition 2}, this is a lower bound on the amount 
bias introduced by the initial tasks and framing.

In experiment \textit{inter-t.1}, the test tasks required workers to
label relatively cluttered images of table settings, which had various items
of food, utensils, dishware, and other objects natural to such settings.
For one treatment, the images in the initial tasks consisted exclusively of 
food, and for the other, they consisted exclusively of non-food objects, 
particularly cookware and utensils.  We saw a strong effect of priming, 
based on the classifier performance (see fig. 2).
